### 2013-04-24 paper/paper.tex
Added paper template and fixed init and readme issues

\documentclass[12pt,a4paper]{report}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\author{Kevin Gao,
Matt Li,
Ashvin Nair, Saavan Patel}
\title{WHS Machine Learning: Lessons Learned}
\begin{document}

\maketitle

For this yearlong autodidactic pioneering computer science course at Winchester High School, we tackled introductory machine learning topics with the intention of understanding the algorithm and implementing them in code. For each topic, we also used the algorithm on real data.

\end{document}